---
title: "2: Central Limit Theorem"
output: powerpoint_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 2.1: Generate Chi-Square Distribution

```{r}
population = rchisq(10000000, 2)
```

## 2.2: Histogram of Chi-Square Distribution

```{r}
hist(population, col = "blue", main = paste("Chi-Square Distribution"), xlab = "Values", xlim = c(0,15))
```

## 2.3: Mean and Standard Deviation of Chi-Square Distribution

### Mean: 

```{r}
mean(population)
```

### Standard Deviation:

```{r}
sd(population)
```

## 2.4: Expected Distribution of Sample Means

### According to the Central Limit Theorem, the distribution of means for samples of size 50 from this population:

- Should be approximately normal
- Should have the same mean as the population
- Should have a standard deviation equal to $\frac{\sigma}{\sqrt{n}}$

## 2.5: Simulated Distribution of Sample Means

```{r}
simulations = 10000
xbar_holder = numeric(simulations)
for (i in 1:simulations)
{
  sample = sample(population, 50)
  xbar = mean(sample)
  xbar_holder[i] = xbar
}
hist(xbar_holder, col = "blue", main = paste("Histogram of Distribution of Sample Means"), xlab = "Sample Mean", xlim = c(0,4))
```

## 2.6: Mean and Standard Deviation of Sample Means

### Mean:

```{r}
mean(xbar_holder)
```

### Standard Deviation:

```{r}
sd(xbar_holder)
```