---
title: "4: T-Test"
output: powerpoint_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Six Steps of Hypothesis Testing

1. Identify the null (H~0~) and alternative (H~a~) hypotheses.
2. Find the critical value.
3. Find the test statistic (t-value) for the data.
4. Find the p-value.
5. Reject or fail to reject H~0~.
6. Present conclusion.

## Step 1: Identify the null (H~0~) and alternative (H~a~) hypotheses.

We are testing the claim that the mean age of the distribution of Beach Comber 
patrons is different than 21. This claim implies that we're testing whether the 
mean could be less than or greater than 21, so we will conduct a two-tailed
t-test:

### H~0~: $\mu = 21$

### H~a~: $\mu \neq 21$

## Step 2: Find the Critical Value

The critical value is found by determining the significance level ($\alpha$) and
degrees of freedom for our test, and looking those values up on a table (or, in
our case, using the qt() function from The R Stats Package).

In the absence of a stated confidence level we assume 95%, which means that our 
significance level is 0.025 (since this is a two-tailed test).

For a two-tailed test, the degrees of freedom for our data is equal to n-2 (in
our case, df = 5).

## Step 2: Code to Determine Critical Value

```{r}
sample = c(25, 19, 37, 29, 40, 28, 31)
significance_level = 0.05/2
degrees_of_freedom = length(sample) - 2
critical_value = qt(p = significance_level, df = degrees_of_freedom)
```

### Critical Value = `r critical_value`

## Steps 3, 4, and 5: Find the Test Statistic, Find the P-Value, and Reject or Fail to Reject H~0~

The t.test() function in the R Stats Package will take our sample vector, with
arguments for the type of test, null hypothesis, and confidence level, and 
return our test statistic, p-value, and a determination as to whether we should
reject or fail to reject H~0~.

## Steps 3, 4, and 5: Code for T-Test

```{r}
t.test(sample, alternative = c("two.sided"), mu = 21, conf.level = 0.95)
```

## Step 6: Present Conclusion

There is enough evidence to suggest that the true mean age of Beach Comber 
patrons is not 21. We are 95% confident that the true mean age of patrons is
between 23.3 and 36.4. However, we encourage the staff to continue to be 
vigilant, as at least one member of our sample was under the age of 21!